Topological order in the quantum dimer-pentamer model

### 2015-09-09 qdpm_paper.tex
Worked on section II

Here we will provide background about generalized quantum dimer models on the square lattice. We consider systems with Ising-like dimer degrees of freedom on each link, such that each link has either $0$ or $1$ dimers. The Hilbert space consists of configurations of dimers on the square lattice that satisfy a local constraint at each vertex; the simplest constraint is that of the hard-core fully-packed QDM where a single dimer must touch each vertex. We write this constraint as $n_v = 1$, where $n_v$ is the number of dimers that touch a vertex. Other unique, non-trivial constraints are $n_v=2$ and the parity constraints $n_v\mod2 =0$ and  $n_v\mod2 =1$ (see \figref{fig:GQDMexamples}).

The Hamiltonian of a GQDM is of the form
\begin{equation}
H_{\rm{GQDM}} = -t \hat{T} + v \hat{V}
\end{equation}
where $\hat{T}$ is an off-diagonal kinetic term that preserves the constraint, $\hat{V}$ is a local potential energy term, and $v$ and $t$ are the strength of each term.

>>>>

Here we will provide background about generalized quantum dimer models on the square lattice. We consider systems with Ising-like dimer degrees of freedom on each link, such that each link has either $0$ or $1$ dimers. The Hilbert space consists of configurations of dimers on the square lattice that satisfy a local constraint at each vertex; the simplest constraint is that of the hard-core fully-packed QDM where a single dimer must touch each vertex. We write this constraint as $n_v = 1$, where $n_v$ is the number of dimers that touch a vertex. We can distinguish between number constraints of the form $n_v = n_0$ and parity constraints of the form $n_v \mod 2 = 0,1$. The unique, non-trivial constraints of this form are the number constraints $n_v=1,2$ and the parity constraints $n_v\mod2 =0,1$ and  $n_v\mod2 =1$ (see \figref{fig:GQDMexamples}).

The presence of such local constraints implies all physical states in the Hilbert space are invariant under local gauge transformations at each vertex, $G_v$.

The Hamiltonian of a GQDM is of the form
\begin{equation}
H_{\rm{GQDM}} = -t \hat{T} + v \hat{V}
\end{equation}
where $\hat{T}$ is an off-diagonal kinetic term that preserves the constraint, $\hat{V}$ is a local potential energy term, and $v$ and $t$ parametrize the strength of each term. We will take the kinetic term $\hat{T}$ to be the dynamics which preserved the constraint; for example in the QDM $\hat{T}$ is the sum of all possible ``plaquette" flip terms acting on plaquettes with parallel dimers.
\begin{equation}
    \label{}
    \hat{T} = \sum_{\bplqt} 
            \left|
                \vrtplqt
            \right\rangle
            \left\langle
                \hrzplqt
            \right|
            +
            \mathrm{h.c.}
\end{equation}
